Comments for the author

### 2021-04-28 Comments_Reviewer.Rmd
Some layout changes
Adding CRS section
Several fixes in the reply to the comments.

**Reply**: Thank you very much for considering the manuscript for publication. 
We appreciate the reviewers comments and think they all help to improve the manuscript.
We agree with reviewer 1 that the definition of landscape ecology could be shortened given the target audience of the journal. 
Thus, the introduction now puts a stronger focus on advantages of open-source software and especially R. 
Furthermore, we added explanations for all packages that were missing in the original submission as suggested by reviewer 1.
However, we think that given recommendations which packages to use is very difficult for several reasons. 
First, packages have different advantages and disadvantages that differ depending on e.g., the research question, the available data, or the personal preferences of the users.
Second, we believe that trying out several packages for similar task can actually beneficial and strengthen trust in the method.
But, by modifying Figure 1 and adding Box 1 we hope to give some more guidelines on how to organize an analysis.
We agree that the survey is not the core of the paper, nevertheless, the comments by both reviewers are very valuable and improve the conclusions we can draw from it. 
Thus, we added some explanation about the survey reach and what this means for the results of the survey. \vskip 0.5cm

@@

**Reply**: We agree that the of Rayfield et al. 2011 is great, but we also want to point out that it was developed for a more specific topic (i.e., connectivity), while this manuscript deals with the whole field of landscape ecology.
Thus, we believe that a similar typology for all packages included in this manuscript would be too complicated to be of actual help.
Thank you for this great idea to add boxes. 
We added a box "Where to start" to give some major steps how to approach landscape ecology in R for novice users. 
We hope this will also at least partly address the issue raised by reviewer 1 about recommendations.
We also thought about a second box for mode advanced users and even developers, but decided that the target audience of the paper are most likely novice to intermediate R users and not developers and the topic of developing own packages specifically for the landscape ecology community could maybe be a future manuscript.
Also, we added a short explanation of the tidyverse and its strength.
Lastly, we provide more information about the annotated references.
\vskip 0.5cm

@@

**Reply**: We hope that the modified Figure 1, the added Box 1 and some added explanation in the conclusions explain how using R can facilitate a reproducible workflow in landcape ecology. \vskip 0.5cm

@@

**Reply**: We agree that it was not clear enough how these package could help woth reproducibility, thus we added a more detailed explanation. \vskip 0.5cm

@@

**Reply**: @Jakub? \vskip 0.5cm

@@

**Reply**: We added that we aware that the survey is most likely not representative for the R community and R users due to how we tried to reach the community and also due to the questions asked in the survey. 
Since the survey is not the main idea of the manuscript though, we think that the survery still migth be interesting even with these shortcomings. \vskip 0.5cm

@@

**Reply**: We agree that the sentence implied that the R community is not constantly adapting to the new developments of landscape ecology which is certainly not true. 
We reworded the sentence making clearea that indeed the community constantly improves the R programming language. 
However, we do not agree with the reviewer that "[...] the problem is that there are so many packages [..]".
Even though we acknowledge that the amount of package can be overwhelming, especially for novices users, diversity of packages is a big strength - even for similar specific tasks. 
Possible advantages include e.g., the possibility to compare different implementations of similar/identical task or providing functionality for slightly different data structures. \vskip 0.5cm

"Because it is possible to import, modify, analyze, and visualize spatial data all in the same programming environment, R allows for transparent and reproducible workflows."
But the authors do not mention any tools for facililitating reproducibility.
This is a huge part of why R should be more widely used, and should be expanded upon.

@@

**Reply**: We strongly agree with the reviewer that this information would be very valuable to judge the representation of the survey.
However, unfortunately such questions were not included.
We also acknowledge that the survey is most likely not representative for all skill levels of *R* users because we assume that by using mailing list and social media more advanced users were more likely to take part in the survey.
We added and clearified those shortcomings of the survey. \vskip 0.5cm

>>>>

**Reply**: Thank you very much for considering the manuscript for publication. 
We appreciate the reviewers comments and think they all help to improve the manuscript.
We agree with reviewer 1 that the definition of landscape ecology could be shortened given the target audience of the journal. 
Thus, the introduction now puts a stronger focus on advantages of open-source software and especially R. 
Furthermore, we added explanations for all packages that were missing in the original submission as suggested by reviewer 1.
However, we think that given recommendations which packages to use is very difficult for several reasons. 
First, packages have different advantages and disadvantages that differ depending on e.g., the research question, the available data, or the personal preferences and experiences of the users.
Second, we believe that trying out several packages for similar task can actually beneficial and strengthen trust in the method.
But, by modifying Figure 1 and adding Box 1 we hope to give some more guidelines on how to organize an analysis.
We agree that the survey is not the core of the paper, nevertheless, the comments by both reviewers are very valuable and improve the conclusions we can draw from it. 
Thus, we added some explanation about the survey reach and what this means for the results of the survey. \vskip 0.5cm

@@

**Reply**: We agree that the approach of Rayfield et al. 2011 is great, but we also want to point out that it was developed for a more specific domain (i.e., connectivity), while this manuscript deals with the whole field of landscape ecology.
Thus, we believe that a similar typology for all packages included in this manuscript would be too complicated to be of actual help.
Thank you for this great idea to add boxes. 
We added a box "Where to start" to give some major steps how to approach landscape ecology in R for novice users. 
We hope this will also at least partly address the issue raised by reviewer 1 about recommendations.
We also thought about a second box for mode advanced users and even developers, but decided that the target audience of the paper are most likely novice to intermediate R users and not developers and the topic of developing own packages specifically for the landscape ecology community could maybe be a future manuscript.
Also, we added a short explanation of the tidyverse and its strength.
Lastly, we provide more information about the annotated references.
\vskip 0.5cm

@@

**Reply**: We hope that the modified Figure 1, the added Box 1 and some added explanation in the conclusions explain how using R can facilitate a reproducible workflow in landscape ecology. \vskip 0.5cm

@@

**Reply**: We agree that it was not clear enough how these package could help with reproducibility, thus we added a more detailed explanation. \vskip 0.5cm

@@

**Reply**: We added an explanation of the recent changes to PROJ and make users aware to pay attention if old Proj4 strings are used. 
Furthermore, we advise to use ESPG codes for CRS. \vskip 0.5cm

@@

**Reply**: We added that we aware that the survey is most likely not representative for the R community and R users due to how we tried to reach the community and also due to the questions asked in the survey. 
Since the survey is not the main idea of the manuscript though, we think that the survey still might be interesting even with these shortcomings. \vskip 0.5cm

@@

**Reply**: We agree that the sentence implied that the R community is not constantly adapting to the new developments of landscape ecology which is certainly not true. 
We reworded the sentence making clearer that indeed the community constantly improves the R programming language. 
However, we do not agree with the reviewer that "[...] the problem is that there are so many packages [..]".
Even though we acknowledge that the amount of package can be overwhelming, especially for novices users, diversity of packages is a big strength - even for similar specific tasks. 
Possible advantages include e.g., the possibility to compare different implementations of similar/identical task or providing functionality for slightly different data structures. \vskip 0.5cm

"Because it is possible to import, modify, analyze, and visualize spatial data all in the same programming environment, R allows for transparent and reproducible workflows."
But the authors do not mention any tools for facilitating reproducibility.
This is a huge part of why R should be more widely used, and should be expanded upon.

@@

**Reply**: We strongly agree with the reviewer that this information would be very valuable to judge the representation of the survey.
However, unfortunately such questions were not included.
We also acknowledge that the survey is most likely not representative for all skill levels of *R* users because we assume that by using mailing list and social media more advanced users were more likely to take part in the survey.
We added and clarified those shortcomings of the survey. \vskip 0.5cm
